### 2021-08-20 doc/thesis/chapters/deployment-on-hardware.tex
unify 'highlights' in the text

The CPU is the main actor that determines any actions on the system.
First, it reads a file from the external SD card and stores this data in a dedicated area in the memory.
Next, the DMA is configured and its data transfer mode is initiated, so that the DMA starts to transfer data from the mentioned memory area towards the FPGA.
The DMA is configured in a mode that creates a continuous transfer, by looping through the data -- once the end is reached, the transfer starts at the beginning of the memory area.
Any data connection in the following processing chain is implemented as an AXI stream bus.
The first block after the DMA is an AXI Stream Switch.
It can switch its single input to be forwarded to either of the outputs.
The switch is configured by the CPU, to enable one output at a time.
Consequently, one of the FM radio receiver IPs is provided with data.
The IP processes this data to produce an audio output, which is delivered to another AXI Stream Switch.
It fulfills the function to either select the output of above or below IP, in coordination with the first switch, obviously.
The next block is an AXI Stream Broadcaster, which takes a single stream input and simply duplicates it at its output, so that the outputs are an exact copy of the input stream.
One of these duplicates is sourced to an AXI-Stream-To-I2S converter, which generates the I2S signals for the external audio codec chip.
The second duplicate is connected to a FIFO, which simply stores the generated output data.
Once the FIFO is almost full, it asserts an interrupt to the CPU, which triggers it to read out the entire FIFO buffer.
The CPU then stores this data on the external SD card again.
In that way, the loop of data is closed, which enables the verification of the respective IPs' correct functionality.
Therefore, this system design can be described as a hardware-in-the-loop system, in the point of view of the FPGA IPs.
This ability refers to the knowledge gained with the SIMILAR concept, that is introduced in Section~\ref{sec:SimilarProcess}.
Specifically, this implemented system design enables the application of the steps 'Assess Performance' and 'Re-evaluate', since the IPs' performance can be measured and their design can be re-evaluated.\\

>>>>

The CPU is the main actor that determines any actions on the system.
First, it reads a file from the external SD card and stores this data in a dedicated area in the memory.
Next, the DMA is configured and its data transfer mode is initiated, so that the DMA starts to transfer data from the mentioned memory area towards the FPGA.
The DMA is configured in a mode that creates a continuous transfer, by looping through the data -- once the end is reached, the transfer starts at the beginning of the memory area.
Any data connection in the following processing chain is implemented as an AXI stream bus.
The first block after the DMA is an AXI Stream Switch.
It can switch its single input to be forwarded to either of the outputs.
The switch is configured by the CPU, to enable one output at a time.
Consequently, one of the FM radio receiver IPs is provided with data.
The IP processes this data to produce an audio output, which is delivered to another AXI Stream Switch.
It fulfills the function to either select the output of above or below IP, in coordination with the first switch, obviously.
The next block is an AXI Stream Broadcaster, which takes a single stream input and simply duplicates it at its output, so that the outputs are an exact copy of the input stream.
One of these duplicates is sourced to an AXI-Stream-To-I2S converter, which generates the I2S signals for the external audio codec chip.
The second duplicate is connected to a FIFO, which simply stores the generated output data.
Once the FIFO is almost full, it asserts an interrupt to the CPU, which triggers it to read out the entire FIFO buffer.
The CPU then stores this data on the external SD card again.
In that way, the loop of data is closed, which enables the verification of the respective IPs' correct functionality.
Therefore, this system design can be described as a hardware-in-the-loop system, in the point of view of the FPGA IPs.
This ability refers to the knowledge gained with the SIMILAR concept, that is introduced in Section~\ref{sec:SimilarProcess}.
Specifically, this implemented system design enables the application of the steps \textit{Assess Performance} and \textit{Re-evaluate}, since the IPs' performance can be measured and their design can be re-evaluated.\\

### 2021-08-20 doc/thesis/chapters/hls.tex
unify 'highlights' in the text

Xilinx is the largest manufacturer of FPGAs globally and therefore its tool 'Xilinx HLS' is widely used.
Intel, the second largest FPGA manufacturer, also has its HLS tool called 'Intel oneAPI' \cite{IntelHLSOneAPI}.
Both companies are conducting serious effort to push the HLS implementation method.
Xilinx recently open-sourced the code for the front-end of their HLS compiler.
The front-end processes the HLS code, specifically \cplusplus, and creates an intermediate representation from it.
The company expects to expand the community engagement and the encouragement of innovation, driven by users, with this step.
Also Intel with its oneAPI toolkit is trying to introduce a wider usage of HLS to the community.\\

In terms of language support, several tools settle for \cplusplus.
Intel oneAPI uses library APIs and DP\cplusplus\ (Data-Parallel \cplusplus), which is an extension to the ISO \cplusplus\ standard.
Xilinx also decided in favor of the \cplusplus\ language for their tools 'Vivado HLS' and 'Vitis HLS'.
In older versions, SystemC was supported for synthesis as well, but is now deprecated in the respective latest versions \cite{VivadoHlsDeprecatesSystemC}.\\

@@

\begin{itemize}
  \item \textbf{Software Testbench}\\
      The developed HLS IP code needs to be verified and checked for correct functionality in a software testbench, before it gets processed to RTL code.
      This can be done by using using conventional software development tools.
      Xilinx specifically recommends dynamic code checker tools like \textit{valgrind} or \textit{Coverity}.
      They support a developer in finding memory leaks, out-of-bounds memory access, uninitialized variables and similar issues.
      Furthermore, the application of code coverage tools like \textit{gcov} is recommended as well.
      In that way, the percentage of executed code lines of testbench and HLS code can be measured, to ensure a sufficient level of testing.
      The software testbench is also called 'C Simulation'.\\

@@

The argumentation is equally valid for integer data types, as well as fixed point data types.
In calculations it is often necessary to represent fractional numbers.
In standard \cplusplus, floating point data types are used therefor.
However, floating point operations require a large amount of logic to implement and thus, HLS provides fixed point data types.
These can be instantiated in arbitrary bitwidths, where integer and fractional bitwidths can be specified separately.
The fixed point format is often described in the format 'integer.fractional', like '2.14', which represents a 16 bit fixed point data type that has 2 integer bits and 14 fractional bits.\\

@@

\begin{itemize}
  \item \textbf{Create input vectors}\\
  In this first step, the above explained, regular C simulation is executed.
  Hereby, the inputs that are sent to the DUT are stored as 'input vectors'.

  \item \textbf{Perform RTL simulation}\\
  Next, the previously generated input vectors are applied to the DUT.
  At this point, the DUT is already translated into RTL code, so the code that is used here is either VHDL or Verilog code.
  The default simulator is the Vivado Simulator (XSim).
  However, various other simulators like ModelSim are supported.
  The generated outputs of this step are stored as 'output vectors'.
  \item \textbf{Verify the results}\\
  The final verification step sources the output vectors back to the C testbench, which compares the outputs with the expected results.\\
\end{itemize}

>>>>

Xilinx is the largest manufacturer of FPGAs globally and therefore its tool \textit{Xilinx HLS} is widely used.
Intel, the second largest FPGA manufacturer, also has its HLS tool called \textit{Intel oneAPI} \cite{IntelHLSOneAPI}.
Both companies are conducting serious effort to push the HLS implementation method.
Xilinx recently open-sourced the code for the front-end of their HLS compiler.
The front-end processes the HLS code, specifically \cplusplus, and creates an intermediate representation from it.
The company expects to expand the community engagement and the encouragement of innovation, driven by users, with this step.
Also Intel with its oneAPI toolkit is trying to introduce a wider usage of HLS to the community.\\

In terms of language support, several tools settle for \cplusplus.
Intel oneAPI uses library APIs and DP\cplusplus\ (Data-Parallel \cplusplus), which is an extension to the ISO \cplusplus\ standard.
Xilinx also decided in favor of the \cplusplus\ language for their tools \textit{Vivado HLS} and \textit{Vitis HLS}.
In older versions, SystemC was supported for synthesis as well, but is now deprecated in the respective latest versions \cite{VivadoHlsDeprecatesSystemC}.\\

@@

\begin{itemize}
  \item \textbf{Software Testbench}\\
      The developed HLS IP code needs to be verified and checked for correct functionality in a software testbench, before it gets processed to RTL code.
      This can be done by using using conventional software development tools.
      Xilinx specifically recommends dynamic code checker tools like \textit{valgrind} or \textit{Coverity}.
      They support a developer in finding memory leaks, out-of-bounds memory access, uninitialized variables and similar issues.
      Furthermore, the application of code coverage tools like \textit{gcov} is recommended as well.
      In that way, the percentage of executed code lines of testbench and HLS code can be measured, to ensure a sufficient level of testing.
      The software testbench is also called \textit{C Simulation}.\\

@@

The argumentation is equally valid for integer data types, as well as fixed point data types.
In calculations it is often necessary to represent fractional numbers.
In standard \cplusplus, floating point data types are used therefor.
However, floating point operations require a large amount of logic to implement and thus, HLS provides fixed point data types.
These can be instantiated in arbitrary bitwidths, where integer and fractional bitwidths can be specified separately.
The fixed point format is often described in the format \textit{integer.fractional}, like \textit{2.14}, which represents a 16 bit fixed point data type that has 2 integer bits and 14 fractional bits.\\

@@

\begin{itemize}
  \item \textbf{Create input vectors}\\
  In this first step, the above explained, regular C simulation is executed.
  Hereby, the inputs that are sent to the DUT are stored as \textit{input vectors}.
  \item \textbf{Perform RTL simulation}\\
  Next, the previously generated input vectors are applied to the DUT.
  At this point, the DUT is already translated into RTL code, so the code that is used here is either VHDL or Verilog code.
  The default simulator is the Vivado Simulator (XSim).
  However, various other simulators like ModelSim are supported.
  The generated outputs of this step are stored as \textit{output vectors}.
  \item \textbf{Verify the results}\\
  The final verification step sources the output vectors back to the C testbench, which compares the outputs with the expected results.\\
\end{itemize}

### 2021-08-20 doc/thesis/chapters/implementation.tex
unify 'highlights' in the text

          To overcome the issue at this point in the actual hardware implementation, the filters were manually 'tuned' to have a matching order.
      \end{itemize}

@@

In the implemented version, an audio file is modulated into the FM channel.
The audio file contains data that represents speech with the words \textit{'Left channel... Right channel...'}.
In terms of audio channels, the words exactly match the respective channels' output -- the first part is modulated on the left audio channel only, while the second part only targets the right audio channel.
This is particulary useful for verification, especially to verify the functionality of channel separation.
From the experience during development, the channel separation is a good indicator of whether the system is functioning correctly.
This left-right indication also simplifies the verification process, because a developer can recognize it with their visual and acoustic senses.
It is possible to audibly verify it by listening to the output.
It is also possible to just visually inspect the output data, where a clear separation is visible between the channels.\\

@@

The Python classes \textit{tb\_analyzer\_helper}, \textit{tb\_data\_handler} and \textit{fm\_receiver\_model} are implemented to create a better structure of the testbench and to provide code that is re-useable for the testbench of the HLS implementation.
The \textit{tb\_analyzer\_helper} class provides functions to analyze the data, such as comparison against defined error thresholds or the creation of graphical plots with Pythons' matplotlib library.
The \textit{tb\_data\_handler} only works as a data storage class, that encapsulates the data that is captured from the DUT.
The class called \textit{fm\_receiver\_model} serves as the 'golden model reference' that is used to compare the DUT against.
Internally, it loads the Matlab verification data.
This could be implemented differently, since Python can be used to actually implement an FM receiver including the entire DSP functionality, just like in Matlab.
Because the Matlab model already exists, this is not done here -- loading the Matlab data is sufficient as it is been seen as 'the single source of truth' in this project.\\

@@

The VHDL code for this register interface is generated by a 'Register Engine' Python script.
The registers can be defined in a YAML-file.
Additionally, documentation in the form of a Markdown language file is generated, and -- more importantly -- a C/\cplusplus\ code header, which defines all the register addresses and can be used in the application software.\\

@@

The testbench is especially designed around the 'Design For Reuse' segment of the SIMILAR concept, which is introduced in Section~\ref{sec:SimilarProcess}.
This is the case because of two major parts.
%It is not only using the exact same Python code for analysis, but is also using parts of the application software code directly, to load the testbench input data.\\

>>>>

          To overcome the issue at this point in the actual hardware implementation, the filters are manually adapted to have a matching order.
      \end{itemize}

@@

In the implemented version, an audio file is modulated into the FM channel.
The audio file contains data that represents speech with the words \textit{Left channel... Right channel...}.
In terms of audio channels, the words exactly match the respective channels' output -- the first part is modulated on the left audio channel only, while the second part only targets the right audio channel.
This is particulary useful for verification, especially to verify the functionality of channel separation.
From the experience during development, the channel separation is a good indicator of whether the system is functioning correctly.
This left-right indication also simplifies the verification process, because a developer can recognize it with their visual and acoustic senses.
It is possible to audibly verify it by listening to the output.
It is also possible to just visually inspect the output data, where a clear separation is visible between the channels.\\

@@

The Python classes \textit{tb\_analyzer\_helper}, \textit{tb\_data\_handler} and \textit{fm\_receiver\_model} are implemented to create a better structure of the testbench and to provide code that is re-useable for the testbench of the HLS implementation.
The \textit{tb\_analyzer\_helper} class provides functions to analyze the data, such as comparison against defined error thresholds or the creation of graphical plots with Pythons' matplotlib library.
The \textit{tb\_data\_handler} only works as a data storage class, that encapsulates the data that is captured from the DUT.
The class called \textit{fm\_receiver\_model} serves as the golden model reference that is used to compare the DUT against.
Internally, it loads the Matlab verification data.
This could be implemented differently, since Python can be used to actually implement an FM receiver including the entire DSP functionality, just like in Matlab.
Because the Matlab model already exists, this is not done here -- loading the Matlab data is sufficient as it is been seen as the single source of truth in this project.\\

@@

The VHDL code for this register interface is generated by a Python script which is named \textit{Register Engine}.
The registers can be defined in a YAML-file.
Additionally, documentation in the form of a Markdown language file is generated, and -- more importantly -- a C/\cplusplus\ code header, which defines all the register addresses and can be used in the application software.\\

@@

The testbench is especially designed around the \textit{Design For Reuse} segment of the SIMILAR concept, which is introduced in Section~\ref{sec:SimilarProcess}.
This is the case because of two major parts.
%It is not only using the exact same Python code for analysis, but is also using parts of the application software code directly, to load the testbench input data.\\

### 2021-08-20 doc/thesis/chapters/results.tex
adapt system-design-process
unify 'highlights' in the text

The cause for this issue may be located in several components in the design.
This includes the FM demodulator, the carrier recovery, i.e. the 38~kHz carrier, as well as a sample timing shift in the final summation and substraction to recover the left and right channels.
The FIR filter has a successfully passing unit test and is thus assumed to be correct.
Also the fixed-point data type with its overflow- and rounding behaviour is suspected to be a potential issue.
Due to the time limitation in the elaboration of this thesis, this issue can not be traced down to the root cause and thus still persists in the current implementation.
However, from the system-level point of view of this thesis, the IP works and can be integrated into the final system.

@@

Here, the results of the simulation testbenches are compared against the values that are read from the FPGA directly.
Ideally, the values should be exactly the same.
However, Figure~\ref{fig:audio_output_compare_tb_vs_hw} shows that this is not always the case.
Again, the HLS implementation does have a matching result, whereas the VHDL variant has deviating values.
The explanation here may be linked with above suspicions, but may also be linked to the initial reset values that exist in the hardware.
In the HLS implementation, the reset logic of several registers is implemented automatically, whereas in VHDL these conditions need to be implemented manually.
Therefore, the IPs internal state, i.e. the register values, may be different at the beginning, which leads to an error propagation throughout the design.
The entire DSP chain is built like a pipeline and therefor it takes a number of samples to process, before the 'wrong' intermediate values are flushed out.
All that may be the cause for the deviating values in the VHDL design.\\

>>>>

The cause for this issue may be located in several components in the design.
This includes the FM demodulator, the carrier recovery, i.e. the 38~kHz carrier, as well as a sample timing shift in the final summation and substraction to recover the left and right channels.
The FIR filter has a successfully passing unit test and is thus assumed to be correct.
Also the fixed point data type with its overflow- and rounding behaviour is suspected to be a potential issue.
Due to the time limitation in the elaboration of this thesis, this issue can not be traced down to the root cause and thus still persists in the current implementation.
However, from the system-level point of view of this thesis, the IP works and can be integrated into the final system.

@@

Here, the results of the simulation testbenches are compared against the values that are read from the FPGA directly.
Ideally, the values should be exactly the same.
However, Figure~\ref{fig:audio_output_compare_tb_vs_hw} shows that this is not always the case.
Again, the HLS implementation does have a matching result, whereas the VHDL variant has deviating values.
The explanation here may be linked with above suspicions, but may also be linked to the initial reset values that exist in the hardware.
In the HLS implementation, the reset logic of several registers is implemented automatically, whereas in VHDL these conditions need to be implemented manually.
Therefore, the IPs internal state, i.e. the register values, may be different at the beginning, which leads to an error propagation throughout the design.
The entire DSP chain is built like a pipeline and therefor it takes a number of samples to process, before the wrong, intermediate values are flushed out.
All that may be the cause for the deviating values in the VHDL design.\\

### 2021-08-20 doc/thesis/chapters/system-architecture-and-concept.tex
unify 'highlights' in the text

An example of a system concept and its architecture is developed and shown in Figure~\ref{fig:bd_system_overview_detailed}.
Here, a common data source is provided by the block called \textit{Front End}.
It records data from an antenna and saves it to a file, in a format that the consequent blocks require at their input.
Therefore, a reproducable input is ensured.
Within the \textit{FM Radio Receiver} block, all the different implementation variants are shown in parallel.
They all work on this common input source and generate their respective output.
This output data is again written to a file.
Hereby it is important to note, that the output of the \textit{Matlab} implementation is chosen to be the 'golden reference model' of correct output.
Its output is used as the verification data later in the process.
This is because Matlab was used to research decoding algorithms and to develop the DSP chain, which is used in the GNU Radio, HLS and VHDL implementations.
In addition to the verification data, the Matlab model also writes common parameters to a file.
These parameters are directly used in the other implementations, to ensure a common setup.
The final block is the \textit{Verification} entity, which compares the output files of all the variants, with respect to the chosen model parameters.\\

>>>>

An example of a system concept and its architecture is developed and shown in Figure~\ref{fig:bd_system_overview_detailed}.
Here, a common data source is provided by the block called \textit{Front End}.
It records data from an antenna and saves it to a file, in a format that the consequent blocks require at their input.
Therefore, a reproducable input is ensured.
Within the \textit{FM Radio Receiver} block, all the different implementation variants are shown in parallel.
They all work on this common input source and generate their respective output.
This output data is again written to a file.
Hereby it is important to note, that the output of the \textit{Matlab} implementation is chosen to be the \textit{golden reference model} of correct output.
Its output is used as the verification data later in the process.
This is because Matlab was used to research decoding algorithms and to develop the DSP chain, which is used in the GNU Radio, HLS and VHDL implementations.
In addition to the verification data, the Matlab model also writes common parameters to a file.
These parameters are directly used in the other implementations, to ensure a common setup.
The final block is the \textit{Verification} entity, which compares the output files of all the variants, with respect to the chosen model parameters.\\

### 2021-08-20 doc/thesis/chapters/system-design-process.tex
unify 'highlights' in the text

In this thesis, only the HLS of plain \cplusplus\ is covered.
SystemC is not discussed, since it may potentially be less important in the future.
In the latest Xilinx HLS tools 'Vitis HLS' and 'Vivado HLS', the language SystemC is already deprecated \cite{VivadoHlsDeprecatesSystemC}.\\

@@

The implementation of low-level HDL in the register transfer level (RTL) is denoted as 'direct', or 'manual' implementation in this thesis.\\

>>>>

In this thesis, only the HLS of plain \cplusplus\ is covered.
SystemC is not discussed, since it may potentially be less important in the future.
In the latest Xilinx HLS tools \textit{Vitis HLS} and \textit{Vivado HLS}, the language SystemC is already deprecated \cite{VivadoHlsDeprecatesSystemC}.\\

@@

The implementation of low-level HDL in the register transfer level (RTL) is denoted as \textit{direct}, or \textit{manual} implementation in this thesis.\\
